Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-09-13 fully-abstract-comp/technical-report.tex
Removed the local capabilities (done quick - there may be things I missed).

% Two arguments
\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
\newcommand{\sisptr}[2]{\twoinstr{\src{gettype}}{#1}{#2}}
\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
\newcommand{\sgetloc}[2]{\twoinstr{\src{getloc}}{#1}{#2}}
\newcommand{\sgetlin}[2]{\twoinstr{\src{getlin}}{#1}{#2}}
\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
\newcommand{\ssplice}[2]{\twoinstr{\src{splice}}{#1}{#2}}
\newcommand{\ssplit}[2]{\twoinstr{\src{split}}{#1}{#2}}

@@

% Two arguments
\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
\newcommand{\tisptr}[2]{\twoinstr{\trg{gettype}}{#1}{#2}}
\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
\newcommand{\tgetloc}[2]{\twoinstr{\trg{getloc}}{#1}{#2}}
\newcommand{\tgetlin}[2]{\twoinstr{\trg{getlin}}{#1}{#2}}
\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}
\newcommand{\tsplice}[2]{\twoinstr{\trg{splice}}{#1}{#2}}
\newcommand{\tsplit}[2]{\twoinstr{\trg{split}}{#1}{#2}}

@@

%%% Source dom
\newcommand{\shareddom}[1]{\mathrm{#1}}
\newcommand{\RegName}{\shareddom{RegisterName}}
\newcommand{\Addr}{\shareddom{Addr}}
\newcommand{\Seal}{\shareddom{Seal}}
\newcommand{\Perm}{\shareddom{Perm}}
\newcommand{\Caps}{\shareddom{Cap}}
\newcommand{\SealableCaps}{\shareddom{SealableCap}}
\newcommand{\Word}{\shareddom{Word}}
\newcommand{\Mem}{\shareddom{Memory}}
\newcommand{\Reg}{\shareddom{RegisterFile}}
\newcommand{\Stk}{\shareddom{Stack}}
\newcommand{\Conf}{\shareddom{Conf}}
\newcommand{\ExecConf}{\shareddom{ExecConf}}
\newcommand{\Global}{\shareddom{Global}}
\newcommand{\Linear}{\shareddom{Linear}}
\newcommand{\MemSeg}{\shareddom{MemorySegment}}
\newcommand{\StkFrame}{\shareddom{StackFrame}}
\newcommand{\Stack}{\shareddom{Stack}}

\newcommand{\scbnf}{\shareddom{sc}}
\newcommand{\cbnf}{\shareddom{c}}
\newcommand{\permbnf}{\shareddom{perm}}
\newcommand{\addrbnf}{\shareddom{a}}
\newcommand{\basebnf}{\shareddom{base}}
\newcommand{\aendbnf}{\shareddom{end}}
\newcommand{\glbnf}{\shareddom{gl}}
\newcommand{\linbnf}{\shareddom{l}}
\newcommand{\sealbasebnf}{\sigma_\shareddom{base}}
\newcommand{\sealendbnf}{\sigma_\shareddom{end}}

\newcommand{\sstk}{\shareddom{stk}}
\newcommand{\smsstk}{\shareddom{ms_{stk}}}
\newcommand{\sstkframe}{\shareddom{frame}}
\newcommand{\sopc}{\shareddom{opc}}
\newcommand{\sastk}{\shareddom{a_{stk}}}
\newcommand{\perm}{\var{perm}}
\newcommand{\gl}{\var{g}}
\newcommand{\lin}{\var{l}}
\newcommand{\base}{\shareddom{base}}
\newcommand{\aend}{\shareddom{end}}
\newcommand{\addr}{\shareddom{a}}
\newcommand{\scap}{\shareddom{c}}
\newcommand{\sms}{\shareddom{ms}}

@@

%%% locality
\newcommand{\plainlocality}[1]{\mathrm{#1}}
\newcommand{\glob}{\plainlocality{global}}
\newcommand{\local}{\plainlocality{local}}

@@

\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
\newcommand{\encPerm}[1]{\plainfun{encocePermission}{#1}}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}

@@

\dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
\lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\glbnf & \defbnf & \glob \mid \local \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\glbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\glbnf$ defines domain $\Global$
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

\subsection{Syntax}
\label{sec:syntax}
The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetloc{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
      & & \tsplice{\trg{r}}{\trg{r}} \mid  \tsplit{\trg{r}}{\trg{r}}
\end{array}
\]

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{figure}
\lau{30-08-2017: We should maybe add a picture with local/global hieararchy.}
\lau{01-09-2017: For now, I will add the linear capabilities such that a linear capability stays linear. It does, however, seem like it would be okay to have linear capabilities become non-linear (I at least don't see anything that would break down completely). One draw back would be that one could ``ruin'' the stack capability by making it non-linear.}
\dominique{7-9-2017: Err... you mean that we could allow non-linear capabilities to be made linear, right?  We definitely do not want to allow the stack or return capability to be made non-linear, as we want to prevent it from being aliased.}
\lau{07-09-2017: I actually did mean linear to non-linear (non-linear to linear does not make sense because the non-linear capability may have aliases). The operation that makes a linear capability non-linear would have to clear said capability. If a linear capability has been made non-linear, then someone who expects a linear capability can observe this difference and behave accordingly, i.e., fail the execution. It will, however, remain an invariant that there are no aliases for linear capabilities.}
\lau{I had forgotten that return capabilities probably need to be linear. This would call for a ``load from offset'' operation to be able to load an in code stored seal (see the figure I drew about linking/seals).}

@@

\[
  \writeLocalAllowed{} \defeq \dots
\]

\[
  % Lau: this should work for everything that might be local - used in store.
  \isLoc{c} \defeq \dots
\]

@@

\[
  \nonLoc{c} \defeq \dots
\]

@@

\subsubsection{Source Language}
\label{sec:source-language}
\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated. It was $\sxjmp{}{}$ that needed an update.}

\textbf{Call}\\
Call has three cases, a successful call, undefined and failed. For a call to be successful, it needs two sealed capabilities - a data and a code capability. The code capability must be non-executable and the stack register must contain a stack pointer. Futher the code and data capability must be local and there must be no attempt at passing on the stack capability.
\lau{30-08-2017: Add more description here.}

For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee or the called capabilities must be non-local \lau{30-08-2017: I have forgotten why this is an issue, but I think it is because we are crossing a security domain?}
\dominique{7-9-2017: We declare the call to be undefined if the caller is doing something ``stupid'' that would break security but which we do not want to check for.  This includes: leaving a copy of the stack pointer or the old return pointer in a place accessible to the callee, invoking local callbacks (see discussion in POPL2018 submission of why this is bad).  The condition that $\cb_2 \neq \stkptr{\_,\_,\_,\_}$ is weird because this would not be the caller's fault, but the callee's.  Further thought needed regarding correct blame assignment for this case.}

>>>>

% Two arguments
\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
\newcommand{\sisptr}[2]{\twoinstr{\src{gettype}}{#1}{#2}}
\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
%\newcommand{\sgetloc}[2]{\twoinstr{\src{getloc}}{#1}{#2}}
\newcommand{\sgetlin}[2]{\twoinstr{\src{get}}{#1}{#2}}
\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
\newcommand{\ssplice}[2]{\twoinstr{\src{splice}}{#1}{#2}}
\newcommand{\ssplit}[2]{\twoinstr{\src{split}}{#1}{#2}}

@@

% Two arguments
\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
\newcommand{\tisptr}[2]{\twoinstr{\trg{gettype}}{#1}{#2}}
\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
%\newcommand{\tgetloc}[2]{\twoinstr{\trg{getloc}}{#1}{#2}}
\newcommand{\tgetlin}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}
\newcommand{\tsplice}[2]{\twoinstr{\trg{splice}}{#1}{#2}}
\newcommand{\tsplit}[2]{\twoinstr{\trg{split}}{#1}{#2}}

@@

%%% Source dom
\newcommand{\shareddom}[1]{\mathrm{#1}}
\newcommand{\RegName}{\shareddom{RegisterName}}
\newcommand{\Addr}{\shareddom{Addr}}
\newcommand{\Seal}{\shareddom{Seal}}
\newcommand{\Perm}{\shareddom{Perm}}
\newcommand{\Caps}{\shareddom{Cap}}
\newcommand{\SealableCaps}{\shareddom{SealableCap}}
\newcommand{\Word}{\shareddom{Word}}
\newcommand{\Mem}{\shareddom{Memory}}
\newcommand{\Reg}{\shareddom{RegisterFile}}
\newcommand{\Stk}{\shareddom{Stack}}
\newcommand{\Conf}{\shareddom{Conf}}
\newcommand{\ExecConf}{\shareddom{ExecConf}}
%\newcommand{\Global}{\shareddom{Global}}
\newcommand{\Linear}{\shareddom{Linear}}
\newcommand{\MemSeg}{\shareddom{MemorySegment}}
\newcommand{\StkFrame}{\shareddom{StackFrame}}
\newcommand{\Stack}{\shareddom{Stack}}

\newcommand{\scbnf}{\shareddom{sc}}
\newcommand{\cbnf}{\shareddom{c}}
\newcommand{\permbnf}{\shareddom{perm}}
\newcommand{\addrbnf}{\shareddom{a}}
\newcommand{\basebnf}{\shareddom{base}}
\newcommand{\aendbnf}{\shareddom{end}}
%\newcommand{\glbnf}{\shareddom{gl}}
\newcommand{\linbnf}{\shareddom{l}}
\newcommand{\sealbasebnf}{\sigma_\shareddom{base}}
\newcommand{\sealendbnf}{\sigma_\shareddom{end}}

\newcommand{\sstk}{\shareddom{stk}}
\newcommand{\smsstk}{\shareddom{ms_{stk}}}
\newcommand{\sstkframe}{\shareddom{frame}}
\newcommand{\sopc}{\shareddom{opc}}
\newcommand{\sastk}{\shareddom{a_{stk}}}
\newcommand{\perm}{\var{perm}}
%\newcommand{\gl}{\var{g}}
\newcommand{\lin}{\var{l}}
\newcommand{\base}{\shareddom{base}}
\newcommand{\aend}{\shareddom{end}}
\newcommand{\addr}{\shareddom{a}}
\newcommand{\scap}{\shareddom{c}}
\newcommand{\sms}{\shareddom{ms}}

@@

%%% locality
%\newcommand{\plainlocality}[1]{\mathrm{#1}}
%\newcommand{\glob}{\plainlocality{global}}
%\newcommand{\local}{\plainlocality{local}}

@@

\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
\newcommand{\encPerm}[1]{\plainfun{encocePermission}{#1}}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}

@@

\dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
\lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
\lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
\lau{13-09-2017: For now I have added linearity to the seals.}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

\subsection{Syntax}
\label{sec:syntax}
The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
      & & \tsplice{\trg{r}}{\trg{r}} \mid  \tsplit{\trg{r}}{\trg{r}}
\end{array}
\]

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{figure}
\lau{01-09-2017: For now, I will add the linear capabilities such that a linear capability stays linear. It does, however, seem like it would be okay to have linear capabilities become non-linear (I at least don't see anything that would break down completely). One draw back would be that one could ``ruin'' the stack capability by making it non-linear.}
\dominique{7-9-2017: Err... you mean that we could allow non-linear capabilities to be made linear, right?  We definitely do not want to allow the stack or return capability to be made non-linear, as we want to prevent it from being aliased.}
\lau{07-09-2017: I actually did mean linear to non-linear (non-linear to linear does not make sense because the non-linear capability may have aliases). The operation that makes a linear capability non-linear would have to clear said capability. If a linear capability has been made non-linear, then someone who expects a linear capability can observe this difference and behave accordingly, i.e., fail the execution. It will, however, remain an invariant that there are no aliases for linear capabilities.}
\lau{07-09-2017: I had forgotten that return capabilities probably need to be linear. This would call for a ``load from offset'' operation to be able to load an in code stored seal (see the figure I drew about linking/seals).}
\lau{13-09-2017: Note that most of the above discussion is now captured in Section~\ref{sec:linear-cap}}

@@

@@

@@

\subsubsection{Source Language}
\label{sec:source-language}
\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated (It was $\sxjmp{}{}$ that needed an update.}

\textbf{Call}\\
Call has three cases, a successful call, undefined and failed. For a call to be successful, it needs two sealed capabilities - a data and a code capability. The code capability must be non-executable and the stack register must contain a stack pointer.
\lau{30-08-2017: Add more description here.}

For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee,
\dominique{7-9-2017: We declare the call to be undefined if the caller is doing something ``stupid'' that would break security but which we do not want to check for.  This includes: leaving a copy of the stack pointer or the old return pointer in a place accessible to the callee, invoking local callbacks (see discussion in POPL2018 submission of why this is bad).  The condition that $\cb_2 \neq \stkptr{\_,\_,\_,\_}$ is weird because this would not be the caller's fault, but the callee's.  Further thought needed regarding correct blame assignment for this case.}

If we are not in one of the two cases above, then the execution fails.
\begin{align*}
  \ssem{\scall{r_1}{r_2}} = & 
    \begin{cases}
      (\mem,\reg',\stk',\ms_\stk') & \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        % "red conditions" (in reference to blackboard picture)
                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
                                        \nonExec{cb_2} \text{ and } \\
                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
                                        % "yellow conditions"
                                        \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
                                        \noRetStkReg{\reg} \text{ and } \\
                                        \noRetStkMs{\ms_\stk'} \text{ and } \\
                                        \reg' = \reg
                                        \begin{array}[t]{l}
                                        \update{\pcreg}{\cb_1}\update{\rdata}{\cb_2}\\
                                        \update{\rretc}{\retptr}\\
                                        \update{\rstk}{\stkptr{\perm,[\aaddr,\eaddr],\aaddr}} \text{ and }
                                        \end{array}
\\                                        \text{ for $i \in \{1,2\}$ if $\isLinear{\Phi(r_i)}$, then $\reg''(r_i) = 0$ and} \\
                                        \opc = \Phi(\pcreg)\update{a}{\Phi(\pcreg).a+\calllen} \text{ and } \\
                                        \ms_\priv = \ms_\stk \setminus [\aaddr,\eaddr] \text{ and } \\
                                        \stk' = (\opc,\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
                                        \ms_\stk' = \ms_\stk |_{[\baddr,\eaddr]}
                                      \end{array}\\
                                      & \\
      \tundefined & \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        % "red conditions" (reference to blackboard picture)
                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
                                        \nonExec{cb_2} \text{ and } \\
                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
                                        % "yellow conditions"
                                        \text{not } \arraycolsep=0pt
                                        \left(\begin{array}{l}
                                          \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
                                          \noRetStkReg{\reg} \text{ and } \\
                                          \noRetStkMs{\ms_\stk'}
                                        \end{array} \right)
                                      \end{array}\\
                                      & \\
      \failed & \text{otherwise}
    \end{cases}
\end{align*}

The $\scseal{}{}{}$ instruction seals a sealable capability with a seal. The base of the seal is used for the seal. The used seal only works if it still grants authority to seal (that is the base of the seal is smaller than or equal to the end).
\begin{align*}
  \ssem{\scseal{r_1}{r_2}{r_3}} = & 
                                    \begin{cases}
                                      \Phi\updReg{r_1}{\vsc} & 
                                      \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        \Phi(r_2) = c \in \SealableCaps \text{ and } \\
                                        \Phi(r_3) = \seal{\lin,\sigma,\_} \text{ and } \\
                                        \vsc = \sealed{c,\sigma}
                                      \end{array}\\
                                      & \\
                                      \failed & \text{otherwise}
                                    \end{cases}
\end{align*}
